---
title: "WUS SWE Reanalysis Visualization"
author: "Jack Tarricone"
date: "2022-09-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Visualizing snow accumulation and ablation

-Creating annual snow accumulation and ablation gif -Western United
States UCLA Daily Snow Reanalysis, Version 1 (WUS_UCLA_SR) -URL:
<https://nsidc.org/data/wus_ucla_sr/versions/1#anchor-1> -Data
Contributor(s): Fang, Y., Y. Liu, and S. A. Margulis., UCLA -By: Jack
Tarricone

### Load packages

```{r, load-packages, message = FALSE}
library(terra) # foundational package for raster and vector spatial data analysis
library(lubridate) # manipulating/formatting dates
library(ncdf4) # working with netcdf data
library(ggmap) # making maps
library(gganimate) # gif making
library(ggplot2);theme_set(theme_minimal(12)) # plotting
library(RColorBrewer) # color scales
```

### Data download

-programmatically downloading data from NSIDC using wget -for
reproducibility i'm showing this in R, but i would normally do it in the
terminal -you also but downloaded manually here:
-<https://nsidc.org/data/wus_ucla_sr/versions/1#anchor-1>

```{r}
# grab the wy2018 tile link from nsidc
download_link <-"https://n5eil01u.ecs.nsidc.org/SNOWEX/WUS_UCLA_SR.001/2018.10.01/WUS_UCLA_SR_v01_N39_0W121_0_agg_16_WY2018_19_SWE_SCA_POST.nc"

# make string for wget download
wget_code <-paste0("wget --load-cookies ~/.urs_cookies --save-cookies ~/.urs_cookies --keep-session-cookies --no-check-certificate --auth-no-challenge=on -r --reject 'index.html*' -np -e robots=off ", download_link)
print(wget_code)

# download using wget to same folder as markdown script
# system(wget_code)
```

```{r}
# netcdf file path
wy2018 <-list.files(pattern = "SWE_SCA_POST.nc$", full.names = TRUE)
wy2018 # check
```

```{r}
# read in netcdf
ncin <- nc_open(wy2018) # open tahoe tile
ncin # inspect

# convert the SWE variable to an array
swe_array <- ncvar_get(ncin,"SWE_Post") # read in

# pull out med SWE from third layer
med_swe_array <-swe_array[,,3,] 
```

```{r}
# convert to raster stack and project
stack <-rast(med_swe_array)
ext(stack)<- ext(-121,-120,39,40)
crs(stack) <-"epsg:4326"
stack # inspect
plot(stack[[200]])
```

### Preparing data for plotting

While `terra::plot` is suitable for quick visualizations, it doesn't
have functionality to make publication quality figures. For this,
`ggplot` and `geom_raster` are much better suited. To use `geom_raster`,
`SpatRaster` data needs to be converted into a data frame using the
`as.data.frame` function.

```{r}
# convert to df for ggplot
swe_df <-as.data.frame(stack, xy = TRUE) # xy = TRUE creates lat and lon columns
dim(swe_df) # 50655 rows, by 367 columns (1 lat, 1 lon, 365 days)

## rename columns to dates for plotting
# create dates sequence
dates_seq <-seq(ymd(origin = "2017-10-01"), ymd("2018-09-30"),1)
colnames(swe_df)[3:367] <-as.character(dates_seq) # rename cols to date, leave at  character string
swe_df[swe_df == 0] <-NA # convert 0 to NA for plotting
```

### Using `ggmap` to import a basemap

```{r, message = FALSE}
# set map bounding box
loc <-c(-121,39,-120,40)

# download google sat
my_map <- get_map(location = loc,
                 source = "google", 
                 maptype = "satellite", 
                 crop = TRUE)

# see map, looks good
ggmap(my_map)
```

### Plot using `geom_raster`

```{r}
# define color scale
blues_scale <-brewer.pal(9, 'Blues')

# plot april 1st, 2018
ggmap(my_map) +
  geom_raster(swe_df, # first 
              mapping = aes(x,y, fill = `2018-04-01`)) +
  coord_equal()+
  scale_fill_gradientn(colours = blues_scale, 
                       limits = c(0,2.5), # set color scale bounds
                       na.value="transparent") + # makes no snow values transparent
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + # black border
  labs(title = "North Lake Tahoe SWE: 2018-04-01", 
       fill = "SWE (m)", # legend title
       x="Lon (deg)", y="Lat (deg)")
```

### Make a gif using ```gganimate```
```{r}
# pivot for making gif
swe_df_gif <-tidyr::pivot_longer(swe_df,3:367) # pivot
colnames(swe_df_gif)[3:4] <-c("date","swe") # rename
swe_df_gif$date <-ymd(swe_df_gif$date) # convert to date from character

# plot for animation
plot <-ggmap(my_map) +
  geom_raster(swe_df_gif, 
              mapping = aes(x,y, 
                            group = date, # for the gif
                            fill = swe)) + # what data is being displayed
  coord_equal() + # set aspect ratio
  scale_fill_gradientn(colours = blues_scale, 
                       limits = c(0,2.5), 
                       na.value="transparent") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  labs(title = "Northern Sierra SWE: {frame_time}", # {frame_time} will be date values
       fill = "SWE (m)",
       x="Lon (deg)", y="Lat (deg)")+
  transition_time(date) + # format gif by date
  ease_aes("linear") # how transitions happen
```